---
title: "Week4 Project1 - ChessPlayerStats"
author: "humbertohp"
date: "September 23, 2018"
output:
  html_document:
    highlight: pygments
    theme: cerulean
---

# Load data set from Github repo
```{r, echo=TRUE}

chess_raw <- read.delim("https://raw.githubusercontent.com/humbertohpgit/MSDS1stSem/master/chess_raw.txt", header = FALSE, sep = "\t", stringsAsFactors=FALSE)
chess_raw
is(chess_raw)

library(stringr)

```

# Extract & format Full Names
```{r, echo=TRUE}
playernames <- data.frame(str_extract_all(chess_raw, "[[:upper:]]{3,} [[:upper:]]{1,} [[:upper:]]{0,}[ -][[:upper:]]{0,}"), stringsAsFactors=FALSE)
playernamesfmt <- data.frame(lapply(playernames, function(x) paste(x, ",")), stringsAsFactors=FALSE)
names(playernamesfmt) <- "PlayerName"
playernamesfmt
```


# Extract & format States
```{r, echo=TRUE}
playerstates <- str_extract_all(chess_raw, "[[:upper:]]{2} \\|")
playerstatesfmt <- data.frame(str_extract_all(playerstates, "[[:alpha:]]{2}"), stringsAsFactors=FALSE)
playerstatesfmt2 <- data.frame(lapply(playerstatesfmt, function(x) paste(x, ",")), stringsAsFactors=FALSE)
names(playerstatesfmt2) <- "PlayerState"
playerstatesfmt2
```

# Extract & format Total Points
```{r, echo=TRUE}
playerpoints <- data.frame(str_extract_all(chess_raw, "[0-9]\\.[0-9]"), stringsAsFactors=FALSE)
playerpointsfmt <- data.frame(lapply(playerpoints, function(x) paste(x, ",")), stringsAsFactors=FALSE)
names(playerpointsfmt) <- "PlayerPoints"
playerpointsfmt
```

# Extract & format Pre-Rating
```{r, echo=TRUE}
playerprerating <- str_extract_all(chess_raw, "R:[[:space:]]{1,}[[:alnum:]]{3,}[[:space:]]{0,}")
playerpreratingfmt <- data.frame(str_extract_all(playerprerating, " [[:alnum:]]{3,}"), stringsAsFactors=FALSE)
playerpreratingfmt2 <- data.frame(lapply(playerpreratingfmt, function(x) paste(x, ",")), stringsAsFactors=FALSE)
names(playerpreratingfmt2) <- "PlayerPreRating"
playerpreratingfmt2
```

# Extract and Calculate Avg. Pre Rating of Opponents
## Extract Opponents for each Player
```{r, echo=TRUE}
opps <- str_extract_all(chess_raw, "\\|[1-6]\\.[0-5]  \\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|")
opps2 <- str_extract_all(opps, "\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|[WDLHUBX][[:print:]]{1,4}\\|")
opps2 <- data.frame(opps2, stringsAsFactors=FALSE)
names(opps2) <- "Opps"
opps3 <- lapply(opps2, function(x) str_extract_all(x, "[[:digit:]]{1,2}"))
opps3
```
## Build table of PreRatings for each Player/Opponent
```{r, echo=TRUE}
plyrpreratcalc <- data.frame(str_extract_all(playerprerating, " [[:digit:]]{3,}"), stringsAsFactors=FALSE)
plyrpreratcalcfmt <- data.frame(cbind(c(1:64), plyrpreratcalc))
names(plyrpreratcalcfmt) <- c("Player", "PlayerPreRating")
plyrpreratcalcfmt$PlayerPreRating <- as.numeric(plyrpreratcalcfmt$PlayerPreRating)
plyrpreratcalcfmt
```
## Build table of Opponents for each Player
```{r, echo=TRUE}
opps5 <- c(0,0)
for(i in 1:64){
opps4 <- data.frame(cbind(i, as.numeric(opps3$Opps[[i]])))
opps5 <- rbind(opps5, opps4)
}
names(opps5) <- c("Player", "Opps")
opps5 <- opps5[-1,]
opps5
```
## Map Pre Ratings for each Player's Opponents
```{r, echo=TRUE}
library(dplyr)
opps6 <- opps5 %>% left_join(plyrpreratcalcfmt, by = c("Opps" = "Player"))
opps6
```
## Calculate Avg. Opponent Pre Rating (and give format for final merge)
```{r, echo=TRUE}
oppsavgprerating <- summarise(group_by(opps6,Player), mean(PlayerPreRating))
names(oppsavgprerating) <- c("Player", "OppsAvgPreRating")
oppsavgpreratingfmt <- oppsavgprerating[,-1]
oppsavgpreratingfmt
```

# Build the final dataframe including all the data points (columns) calculated: PlayerName, PlayerState, PlayerPoints, PlayerPreRating, OppsAvgPreRating
```{r, echo=TRUE}
player_df <- cbind(playernamesfmt, playerstatesfmt2, playerpointsfmt, playerpreratingfmt2, oppsavgpreratingfmt)
str(player_df)
head(player_df,1)
player_df
```

# Write the final result set to a CSV file
```{r, echo=TRUE}
setwd("C:/DATA/HHP/Personal/Degrees/Ms. Data Science (CUNY)/R Working Dir")
write.csv(player_df, "ChessPlayerStats.csv", row.names = TRUE)
```



