---
title: "Week9 Assignment - Web APIs (Dashboard style and tabsets)"
author: "humbertohp"
date: "October 27, 2018"
output:
  flexdashboard::flex_dashboard
---
    
#### WEB APIs
#### The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs. Sign up for an API key.
#### Task: Choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe. 
#### Books API - Provides information about books reviews and the New York Times bestsellers lists
   
Column {.tabset}
-------------------------------------
   
### NYT Top 5 Bestsellers for a specific month - October, 2018 (Using jsonlite)

```{r, echo=TRUE}
library(jsonlite)
library(knitr)
api_key <- "&api-key=9507c7df7afe48a08c9d07886b7f97d4"
url_j <- "http://api.nytimes.com/svc/books/v2/lists/overview.json?"
pub_date <- "published_date=2018-10-01"
request_j <- fromJSON(paste0(url_j, pub_date, api_key))
bsellers_lists <- request_j$results$list
bsellers_books <- bsellers_lists[[1, "books"]]
bsellers_df <- subset(bsellers_books, select = c("title", "author", "description", "rank"))
class(bsellers_df)
kable(bsellers_df,  format = "html", caption = "NYT Top 5 Bestsellers - October 2018")
```
 
### Bestsellers History for a specific author: Dan Brown (Using httr)

```{r, echo=TRUE}
library(httr)
library(knitr)
url_h <- "https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json?author=Dan+Brown"
request_h <- GET(url_h, add_headers(api_key="9507c7df7afe48a08c9d07886b7f97d4"))
unlist(http_status(request_h))
bsellers_hist <- content(request_h, "parsed")
num_results <- as.numeric(bsellers_hist$num_results)
bsellers_results <- bsellers_hist$results
bsellers_titles <- data.frame("","")
colnames(bsellers_titles) <- c("Title", "Description")
for (i in 1:num_results){
bsellers_titles_temp <- do.call("cbind", lapply(bsellers_results[[i]][1:2], data.frame, stringsAsFactors = FALSE))
colnames(bsellers_titles_temp) <- c("Title", "Description")
bsellers_titles <- rbind(bsellers_titles, bsellers_titles_temp)}
kable(bsellers_titles,  format = "html", caption = "NYT Bestsellers by Dan Brown")
```


